---
title: "Covid_4"
author: "zly"
date: "2022-11-15"
output: html_document
---
```{r}
library(rehh)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(data.table)
library(DBI)
library(RSQLite)
library(scales)
library(lubridate)
library(data.table)
```

```{r}
filename <- file.choose()
covid_2<- readRDS(filename)
## 少的那个
covid_2 <- covid_2 %>%  filter(grepl(pattern = 'Spike', AASubstitutions))
covid_2 <- covid_2[!is.na(covid_2$date),]
# covid_2 <- df[complete.cases(covid),]
# colnames(covid_2)[names(covid_2) == 'GC-Content'] <- "GCContent"
# saveRDS(covid_2, file = filename)
```


## Week 12
Calculate real dn/ds
```{r}
covid_2$VarShort[is.na(covid_2$VarShort)] <- "Origin"
covid_ns <- covid_2 %>% group_by(pangolin_lineage) %>% arrange(date) %>% mutate(no = row_number(date)) ######### "no" here used to delete 5% data, later will be factor for >1 or not

covid_ns$pangolin_lineage[which(covid_ns$VarShort == "VOC Delta GK")] <- "Delta"


# %>% filter(pangolin_lineage == c("BA.1","BA.2","BA.2.12.1","BA.4","BA.5","BF.7","BQ.1.1"))
covid_ns <- covid_ns[-which(covid_ns$no < 5),]

# covid_2 <- covid_2[order(covid_2$pangolin_lineage,covid_2$date)]

covid_n_1 <- aggregate(covid_ns$count_N, by=list( covid_ns$pangolin_lineage,covid_ns$VarShort,covid_ns$date),mean)

# covid_ns <- data.frame(covid_ns,covid_ns_1)
names(covid_n_1) <- c("pangolin_lineage", "VarShort", "date", "count_N")

covid_n_1 <- covid_n_1[order(covid_n_1$pangolin_lineage,covid_n_1$date),]

# ggplot(covid_n_1, aes(x = date, y = count_N, color = pangolin_lineage)) +
#   geom_point() +
#   labs(title = "Count_N by lineage")

############### N ##############
############### S ##############

covid_s_1 <- aggregate(covid_ns$count_S, by=list(covid_ns$pangolin_lineage,covid_ns$VarShort,covid_ns$date),mean)
# covid_ns <- data.frame(covid_ns,covid_ns_1)
names(covid_s_1) <- c("pangolin_lineage", "VarShort", "date", "count_S")

covid_s_1 <- covid_s_1[order(covid_s_1$pangolin_lineage,covid_s_1$date),]

ggplot(covid_s_1, aes(x = date, y = count_S, color = VarShort)) +
  geom_point() +
  labs(title = "Count_S by lineage")

delta_N <- diff(covid_n_1$count_N)
delta_N <- c(0,delta_N)
covid_n_1$delta_N <- delta_N


delta_S <- diff(covid_s_1$count_S)
delta_S <- c(0,delta_S)
covid_s_1$delta_S <- delta_S

dnds <- data.frame(covid_s_1$date, covid_s_1$pangolin_lineage,delta_N/delta_S)
names(dnds) <- c("date","pangolin_lineage","dnds")


dnds <- dnds[!is.infinite(dnds$dnds),]
dnds <- dnds[!is.na(dnds$dnds),]

# ggplot(dnds, aes(x = date, y = dnds)) +
#   geom_point() +
#   ylim(-100,100) +
#   labs(title = "dNdS")


delta_N <- diff(covid_ns$count_N)
delta_N <- c(0,delta_N)
covid_ns$delta_N <- delta_N


delta_S <- diff(covid_ns$count_S)
delta_S <- c(0,delta_S)
covid_ns$delta_S <- delta_S

covid_ns$dnds <- delta_N/delta_S

covid_ns <- covid_ns[!is.infinite(covid_ns$dnds),]
covid_ns <- covid_ns[!is.na(covid_ns$dnds),]
covid_ns <- covid_ns[!is.na(covid_ns$VarShort),]

# ggplot(covid_ns, aes(x = date, y = dnds, col = pangolin_lineage)) +
#   geom_point()
```

## Week 12

```{r}
############ Linear Regression on N
# lineage_name = c("BA.1","BA.2","BA.2.1","BA.2.75","BA.4","BA.5","BA.5.2.1","BF.7","BQ.1","BQ.1.1","BN.1","XBB")


var_name = c("VOC Omicron GRA", "VOC Delta GK")
# var_name = c("VOC Delta GK")

covid_ns_2 <- covid_ns[covid_ns$VarShort == var_name,]

fitted_N =  covid_ns_2 %>% group_by(pangolin_lineage) %>% do(model = lm(count_N ~ date, data = .))
# plot(fitted_models$model)
# fitted_N$model

######### Linear Regression on S
# covid_s_2 <- covid_ns[covid_ns$VarShort == var_name,]
fitted_S =  covid_ns_2 %>% group_by(pangolin_lineage) %>% do(model = lm(count_S ~ date, data = .))
# plot(fitted_models$model)
# fitted_S$model
# fitted_S <- lm(count_S ~ date,covid_s_2, subset = (grp == "pango"))
# fitted_S
```

```{r}
# ggplot(data = covid_n_2, mapping = aes(x = date, y = count_N, col = covid_n_2$pangolin_lineage)) +
#   # geom_point() +
#   xlab('date') +
#   labs(title = "delta_N(slope)") +
#   geom_smooth(method = lm)
# 
# ggplot(data = covid_n_2, mapping = aes(x = date, y = count_N, col = covid_n_2$pangolin_lineage)) +
#   geom_point() +
#   xlab('date') +
#   labs(title = "count_N") +
#   ylim(0,100)
# 
# ggplot(data = covid_s_2, mapping = aes(x = date, y = count_S, col = covid_s_2$pangolin_lineage)) +
#   # geom_point() +
#   xlab('date') +
#   labs(title = "delta_S(slope)") +
#   geom_smooth(method = lm)
# 
# ggplot(data = covid_s_2, mapping = aes(x = date, y = count_S, col = covid_s_2$pangolin_lineage)) +
#   geom_point() +
#   xlab('date') +
#   labs(title = "count_S") +
#   ylim(0,50)
```


For more lineages
```{r}
a <- list()
for (i in 1:nrow(fitted_N)) {
  a[i] <- fitted_N[[2]][[i]][["coefficients"]][2] / fitted_S[[2]][[i]][["coefficients"]][2]
}  ### bootstrap of dnds
# dnds <- data.frame(matrix(nrow=0,ncol=2))
dnds <- data.frame(fitted_N$pangolin_lineage)
names(dnds) <- c("pangolin_lineage")
dnds$dNdS <- a

# dnds <- dnds[!is.infinite(dnds$dNdS),]
dnds <- dnds[!is.na(dnds$dNdS),]


# dnds <- dnds[!is.na(dnds$dNdS),]

dnds[dnds$pangolin_lineage == "BA.2",]$dNdS
####### Haven't consider p<0.01



# ggplot(data = dnds, aes(x = pangolin_lineage, y = dNdS)) +
#   geom_bar(stat = 'identity')
```


FOR MULTIPLE LINEAGES
```{r}
lineage_name = c("Delta","BA.1","BA.2","BA.2.12.1","BA.2.75","BA.4","BA.5","BA.5.2.1","BF.7","BQ.1","BQ.1.1","BN.1","XBB")
# covid_specific <- covid_2 %>% filter(pangolin_lineage == lineage_name)
# covid_4 <- dnds[dnds$pangolin_lineage == lineage_name,]
dnds_multiple_dnds <- list()
dnds_multiple_lineage <- list()
for (i in 1:length(lineage_name))
{
  dnds_multiple_dnds[i] <- dnds[dnds$pangolin_lineage == lineage_name[i],]$dNdS
  dnds_multiple_lineage[i] <- dnds[dnds$pangolin_lineage == lineage_name[i],]$pangolin_lineage
}
dnds_multiple <- data.frame(dnds_multiple_lineage,dnds_multiple_dnds)

# dnds_multiple <- data.frame(dnds_multiple_lineage)
# names(dnds_multiple) <- c("pangolin_lineage")
# dnds$dNdS <- dnds_multiple_dnds

dnds_multiple <- cbind(dnds_multiple_lineage,dnds_multiple_dnds)
dnds_multiple <- as.data.frame(dnds_multiple)
names(dnds_multiple) <- c("pangolin_lineage","dNdS")
dnds_multiple$dNdS <- as.numeric(dnds_multiple$dNdS)
dnds_multiple$pangolin_lineage <- as.factor(lineage_name)

dnds_multiple$col<- ifelse(dnds_multiple$dNdS >= 1, 1, 0)
dnds_multiple$col<-as.factor(dnds_multiple$col)
dnds_multiple <- dnds_multiple %>%
  mutate(pangolin_lineage = factor(pangolin_lineage, levels = lineage_name))

ggplot(dnds_multiple, mapping = aes(x = dnds_multiple$pangolin_lineage, y = dnds_multiple$dNdS, col = dnds_multiple$col, fill = dnds_multiple$col, levels = lineage_name)) +
  geom_bar(position="dodge",stat="identity") +
  labs(title = "dNdS_new in different Concern Lineages") +
  ylim(0,7)

# dnds[dnds$pangolin_lineage == "XBB",]$dNdS
```
Distribution of a specific lineage
```{r}
# lineage_name = lineage_name
# lineage_name = c("BQ.1")
# covid_specific <- covid_2 %>% filter(pangolin_lineage == lineage_name)

covid_specific <- covid_ns_2[covid_ns_2$pangolin_lineage == lineage_name,]
covid_specific <- covid_specific[!is.na(covid_specific$delta_N),]

dnds[dnds$pangolin_lineage == lineage_name,]$dNdS


ggplot(data = covid_specific, mapping = aes(x = date, y = count_N, col = covid_specific$pangolin_lineage)) +
  # geom_point() +
  xlab('date') +
  labs(title = "delta_N(slope)") +
  geom_smooth(method = lm)

ggplot(data = covid_specific, mapping = aes(x = date, y = count_N, col = covid_specific$pangolin_lineage)) +
  geom_point() +
  xlab('date') +
  labs(title = "count_N")

ggplot(data = covid_specific, mapping = aes(x = date, y = count_S, col = covid_specific$pangolin_lineage)) +
  # geom_point() +
  xlab('date') +
  labs(title = "delta_S(slope)") +
  geom_smooth(method = lm)

ggplot(data = covid_specific, mapping = aes(x = date, y = count_S, col = covid_specific$pangolin_lineage)) +
  geom_point() +
  xlab('date') +
  labs(title = "count_S")

```
BA.2
```{r}

covid_specific_2 <- covid_specific[covid_specific$pangolin_lineage == "BA.2",]
ggplot(data = covid_specific_2, mapping = aes(x = date, y = count_N)) +
  geom_point() +
  xlab('date') +
  labs(title = "count_N BA.2") +
  geom_smooth(method = lm)

covid_specific_2 <- covid_specific[covid_specific$pangolin_lineage == "BA.2",]
ggplot(data = covid_specific_2, mapping = aes(x = date, y = count_S)) +
  geom_point() +
  xlab('date') +
  labs(title = "count_S BA.2") +
  geom_smooth(method = lm)

```

Week 13

Proportion of dnds > 1
```{r}

sum(dnds$dNdS > 1) / nrow(dnds)

dnds$dNdS <-as.numeric(dnds$dNdS)
dnds <- dnds[!is.infinite(dnds$dNdS),]
dnds <- dnds[dnds$dNdS < 10,]
dnds <- dnds[dnds$dNdS > -10,]

# print()mean(dnds$dNdS)
```

Boxplot of dnds by different lineages
```{r}

boxplot(covid_specific$count_N~covid_specific$pangolin_lineage,
  ylim = c(20,60),
  ylab = "delta_N",
  main = "delta_N by lineages")
```

t-test on distribution
```{r}
ks.test(covid_specific$delta_N[covid_specific$VarShort == "VOC Omicron GRA"], covid_specific$delta_N[covid_specific$VarShort == "VOC Delta GK"], 
        paired = F) ### T-test on Delta & Omicron
```
```{r}
ks.test(covid_specific$delta_N[covid_specific$pangolin_lineage == "XBB"], covid_specific$delta_N[covid_specific$pangolin_lineage == "BQ.1.1"], 
        paired = F)


```

Bootstrapping
Delta and Omicron
```{r}

test.stat1 <- mean(covid_specific$count_N[covid_specific$VarShort == "VOC Omicron GRA"]) -
    mean(covid_specific$count_N[covid_specific$VarShort == "VOC Delta GK"])
test.stat1


covid_specific <- covid_specific[covid_specific$VarShort == c("VOC Omicron GRA","VOC Delta GK"),]
# covid_specific$VarShort[covid_specific$`Patient age` > 40] = "VOC Delta GK 2"



set.seed(114514)
n <- length(covid_specific$VarShort)
n

B <- 1000 # num of bs samples
variable <- covid_specific$count_N

BootstrapSample <- matrix(sample(variable, size = n * B, replace = TRUE),
                          nrow = n, ncol = B)
dim(BootstrapSample)

# Init vectors to stor the TEST-STATS
Boot.test.stat1 <- rep(0,B)
Boot.test.stat2 <- rep(0,B)

for (i in 1:B) {
  Boot.test.stat1[i] <- mean(BootstrapSample[1:n/2,i]) -
                              mean(BootstrapSample[floor(n/2+1):n,i])
  
  Boot.test.stat2[i] <- median(BootstrapSample[1:n/2,i]) -
                              median(BootstrapSample[floor(n/2+1):n,i])
}

round(Boot.test.stat1[1:20],1)
mean(Boot.test.stat1 < test.stat1)

plot(density(Boot.test.stat1),
     xlab = expression(bar(Yc) - bar(Ym)),
     main = "mean t-test on Count_N of Delta & Omicron")
```

Targeting on Delta_S
```{r}
test.stat1 <-   mean(covid_specific$count_S[covid_specific$VarShort == "VOC Omicron GRA"]) -
    mean(covid_specific$count_S[covid_specific$VarShort == "VOC Delta GK"])
test.stat1
covid_specific <- covid_specific[covid_specific$VarShort == c("VOC Omicron GRA","VOC Delta GK"),]

set.seed(114514)
n <- length(covid_specific$VarShort)
n

B <- 1000 # num of bs samples
variable <- covid_specific$count_S

BootstrapSample <- matrix(sample(variable, size = n * B, replace = TRUE),
                          nrow = n, ncol = B)
dim(BootstrapSample)

# Init vectors to stor the TEST-STATS
Boot.test.stat1 <- rep(0,B)
Boot.test.stat2 <- rep(0,B)

for (i in 1:B) {
  Boot.test.stat1[i] <- mean(BootstrapSample[1:n/2,i]) -
                              mean(BootstrapSample[floor(n/2+1):n,i])
  
  Boot.test.stat2[i] <- median(BootstrapSample[1:n/2,i]) -
                              median(BootstrapSample[floor(n/2+1):n,i])
}

round(Boot.test.stat1[1:20],1)
mean(Boot.test.stat1 >= test.stat1)

plot(density(Boot.test.stat1),
     xlab = expression(bar(Yc) - bar(Ym)),
     main = "mean t-test on count_S of Delta & Omicron")
```

dnds on important variants
```{r}
# lineage_name = c("Delta","BA.1","BA.2","BA.2.1","BA.2.75","BA.4","BA.5","BA.5.2.1","BF.7","BQ.1","BQ.1.1","BN.1","XBB")
var_name = c("VOI Epsilon GH","VOC Alpha GRY","VOC Gamma GR","VOC Delta GK","VOC Omicron GRA")

# var_name = c("VOC Omicron GRA", "VOC Delta GK")
# var_name = c("VOC Delta GK")

covid_n_2 <- covid_ns[covid_ns$VarShort == var_name,]

fitted_N =  covid_n_2 %>% group_by(VarShort) %>% do(model = lm(count_N ~ date, data = .))
# plot(fitted_models$model)
# fitted_N$model

######### Linear Regression on S
covid_s_2 <- covid_ns[covid_ns$VarShort == var_name,]
fitted_S =  covid_s_2 %>% group_by(VarShort) %>% do(model = lm(count_S ~ date, data = .))
# plot(fitted_models$model)
# fitted_S$model
# fitted_S <- lm(count_S ~ date,covid_s_2, subset = (grp == "pango"))
# fitted_S

covid_n_origin <- covid_n_2 %>% filter(VarShort == "Origin")
ggplot(data = covid_n_origin, mapping = aes(x = date, y = count_N)) +
  geom_point() +
  xlab('date') +
  labs(title = "count_N")
ggplot(data = covid_n_origin, mapping = aes(x = date, y = count_S)) +
  geom_point() +
  xlab('date') +
  labs(title = "count_S")



a <- list()
for (i in 1:nrow(fitted_N)) {
  a[i] <- fitted_N[[2]][[i]][["coefficients"]][2] / fitted_S[[2]][[i]][["coefficients"]][2]
}
# dnds <- data.frame(matrix(nrow=0,ncol=2))
dnds <- data.frame(fitted_N$VarShort)
names(dnds) <- c("VarShort")
dnds$dNdS <- a

# dnds <- dnds[!is.infinite(dnds$dNdS),]
dnds <- dnds[!is.na(dnds$dNdS),]


# dnds <- dnds[!is.na(dnds$dNdS),]

dnds[dnds$VarShort == "VOC Omicron GRA",]$dNdS

# covid_specific <- covid_2 %>% filter(pangolin_lineage == lineage_name)
# covid_4 <- dnds[dnds$pangolin_lineage == lineage_name,]
dnds_multiple_dnds <- list()
dnds_multiple_lineage <- list()
for (i in 1:length(var_name))
{
  dnds_multiple_dnds[i] <- dnds[dnds$VarShort == var_name[i],]$dNdS
  dnds_multiple_lineage[i] <- dnds[dnds$VarShort == var_name[i],]$VarShort
}
dnds_multiple <- data.frame(dnds_multiple_lineage,dnds_multiple_dnds)

# dnds_multiple <- data.frame(dnds_multiple_lineage)
# names(dnds_multiple) <- c("pangolin_lineage")
# dnds$dNdS <- dnds_multiple_dnds

dnds_multiple <- cbind(dnds_multiple_lineage,dnds_multiple_dnds)
dnds_multiple <- as.data.frame(dnds_multiple)
names(dnds_multiple) <- c("VarShort","dNdS")
dnds_multiple$dNdS <- as.numeric(dnds_multiple$dNdS)
dnds_multiple$VarShort <- as.factor(var_name)

dnds_multiple$col<- ifelse(dnds_multiple$dNdS >= 1, 1, 0)
dnds_multiple$col<-as.factor(dnds_multiple$col)
dnds_multiple <- dnds_multiple %>%
  mutate(VarShort = factor(VarShort, levels = var_name))

ggplot(dnds_multiple, mapping = aes(x = dnds_multiple$VarShort, y = dnds_multiple$dNdS, col = dnds_multiple$col, fill = dnds_multiple$col, levels = var_name)) +
  geom_bar(position="dodge",stat="identity") +
  labs(title = "dNdS_new in different Concern Variants")

# dnds[dnds$VarShort == "XBB",]$dNdS
```

Bootstrapping
BQ.1.1 and XBB count_S
```{r}
covid_specific <- covid_n_2[covid_n_2$pangolin_lineage == lineage_name,]
covid_specific <- covid_specific[!is.na(covid_specific$delta_N),]

test.stat1 <- mean(covid_specific$count_S[covid_specific$pangolin_lineage == "BQ.1.1"]) -
    mean(covid_specific$count_S[covid_specific$pangolin_lineage == "XBB"])
test.stat1


covid_specific <- covid_specific[covid_specific$pangolin_lineage == c("XBB","BQ.1"),]
# covid_specific$VarShort[covid_specific$`Patient age` > 40] = "VOC Delta GK 2"



set.seed(114514)
n <- length(covid_specific$pangolin_lineage)
n

B <- 1000 # num of bs samples
variable <- covid_specific$count_S

BootstrapSample <- matrix(sample(variable, size = n * B, replace = TRUE),
                          nrow = n, ncol = B)
dim(BootstrapSample)

# Init vectors to stor the TEST-STATS
Boot.test.stat1 <- rep(0,B)
Boot.test.stat2 <- rep(0,B)

for (i in 1:B) {
  Boot.test.stat1[i] <- mean(BootstrapSample[1:n/2,i]) -
                              mean(BootstrapSample[floor(n/2+1):n,i])
  
  Boot.test.stat2[i] <- median(BootstrapSample[1:n/2,i]) -
                              median(BootstrapSample[floor(n/2+1):n,i])
}

round(Boot.test.stat1[1:20],1)
mean(Boot.test.stat1 < test.stat1)

plot(density(Boot.test.stat1),
     xlab = expression(bar(Yc) - bar(Ym)),
     main = "mean t-test on Count_S of XBB & BQ.1.1")
```

Separate BA.2 by date < Feb.15
```{r}
lineage_name = c("Delta","BA.1","BA.2","BA.2.12.1","BA.2.75","BA.4","BA.5","BA.5.2.1","BF.7","BQ.1","BQ.1.1","BN.1","XBB")
covid_specific_1 <- covid_n_2[covid_n_2$pangolin_lineage == "BA.2",]

ggplot(data = covid_specific_1, mapping = aes(x = date, y = count_N, col = covid_specific_1$pangolin_lineage)) +
  geom_point() +
  xlab('date') +
  labs(title = "count_N") +
  ylim(25,65) +
  geom_smooth(color = "black")

```

```{r}
covid_specific_1$pangolin_lineage[covid_specific_1$date <= "2022-02-15"] = "BA.2A"
covid_specific_1$pangolin_lineage[covid_specific_1$date > "2022-02-15"] = "BA.2B"



test.stat1 <- mean(covid_specific_1$count_N[covid_specific_1$pangolin_lineage == "BA.2A"]) -
    mean(covid_specific_1$count_N[covid_specific_1$pangolin_lineage == "BA.2B"])
test.stat1


covid_specific_1 <- covid_specific_1[covid_specific_1$pangolin_lineage == c("BA.2A","BA.2B"),]
# covid_specific$VarShort[covid_specific$`Patient age` > 40] = "VOC Delta GK 2"



set.seed(114514)
n <- length(covid_specific_1$pangolin_lineage)
n

B <- 1000 # num of bs samples
variable <- covid_specific_1$count_N

BootstrapSample <- matrix(sample(variable, size = n * B, replace = TRUE),
                          nrow = n, ncol = B)
dim(BootstrapSample)

# Init vectors to stor the TEST-STATS
Boot.test.stat1 <- rep(0,B)
Boot.test.stat2 <- rep(0,B)

for (i in 1:B) {
  Boot.test.stat1[i] <- mean(BootstrapSample[1:n/2,i]) -
                              mean(BootstrapSample[floor(n/2+1):n,i])
  
  Boot.test.stat2[i] <- median(BootstrapSample[1:n/2,i]) -
                              median(BootstrapSample[floor(n/2+1):n,i])
}

round(Boot.test.stat1[1:20],1)
mean(Boot.test.stat1 < test.stat1)

plot(density(Boot.test.stat1),
     xlab = expression(bar(Yc) - bar(Ym)),
     main = "mean t-test on Count_N of BA.2A & BA.2B")
```


# Week 14
bootstrap of count_N and count_S   
pd_N pd_S   
```{r}

# library(boot)
set.seed(1234)

covid_specific_1 <- covid_specific[covid_specific$pangolin_lineage == c("BA.2"),]
covid_specific_1 <- covid_specific_1[,c("date","count_N", "count_S")]
pd_NS   <- covid_specific_1

# pd$b <- 0
for (i in 1:5000) {
  j    <- sample(1:nrow(covid_specific_1), replace = TRUE)
  nd   <- covid_specific_1[j,]
  # nd$b <- i
  pd_NS   <- rbind(pd_NS, nd)
  
  if(i%%500==0) {   # Progress tracker
    cat(i, "\n")       
    flush.console()
  }
  
}

ggplot(pd_NS, mapping = aes(x=date, y=count_N)) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  labs(title = "Count_N XBB bs5000")

# covid_specific_1 <- covid_specific[covid_specific$pangolin_lineage == c("BA.2"),]
# covid_specific_1 <- covid_specific_1[,c("date","count_S")]
# pd_S   <- covid_specific_1
# 
# # pd$b <- 0
# for (i in 1:1000) {
#   j    <- sample(1:nrow(covid_specific_1), replace = TRUE)
#   nd   <- covid_specific_1[j,]
#   # nd$b <- i
#   pd_S   <- rbind(pd_S, nd)
#   
# }

ggplot(pd_NS, mapping = aes(x=date, y=count_S)) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  labs(title = "Count_S XBB bs5000")


######################################################  Calculate dnds

fitted_N <- lm(data = pd_NS, count_N ~ date)

######### Linear Regression on S
fitted_S <- lm(data = pd_NS, count_S ~ date)

fitted_N[["coefficients"]][2] / fitted_S[["coefficients"]][2]

```

Calculate dnds
```{r}

fitted_N <- lm(data = pd_NS, count_N ~ date)

######### Linear Regression on S
fitted_S <- lm(data = pd_NS, count_S ~ date)

fitted_N[["coefficients"]][2] / fitted_S[["coefficients"]][2]
```

BA.2A and BA.2B(only when lineage == "BA.2")
```{r}
covid_specific_1 <- covid_specific[covid_specific$pangolin_lineage == c("BA.2"),]
covid_specific_1$pangolin_lineage[covid_specific_1$date <= "2022-02-15"] = "BA.2A"
covid_specific_1$pangolin_lineage[covid_specific_1$date > "2022-02-15"] = "BA.2B"

set.seed(1234)

covid_specific_2 <- covid_specific_1[covid_specific_1$pangolin_lineage == c("BA.2B"),]
covid_specific_2 <- covid_specific_2[,c("date","count_N","count_S")]
pd_NS   <- covid_specific_2

# pd$b <- 0
for (i in 1:2000) {
  j    <- sample(1:nrow(covid_specific_2), replace = TRUE)
  nd   <- covid_specific_2[j,]
  # nd$b <- i
  pd_NS   <- rbind(pd_NS, nd)
  
  if(i%%500==0) {   # Progress tracker
    cat(i, "\n")       
    flush.console()
  }
  
}

ggplot(pd_NS, mapping = aes(x=date, y=count_N)) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  labs(title = "Count_N BA.2B bs2000")
  

ggplot(pd_NS, mapping = aes(x=date, y=count_S)) + 
  geom_point() + 
  geom_smooth(method = "lm")+
  labs(title = "Count_S BA.2B bs2000")

fitted_N <- lm(data = pd_NS, count_N ~ date)

######### Linear Regression on S
fitted_S <- lm(data = pd_NS, count_S ~ date)

fitted_N[["coefficients"]][2] / fitted_S[["coefficients"]][2]
```

```{r}

ggplot(covid_specific_1, aes(x = date, y = count_S, col = pangolin_lineage)) +
  geom_point() +
  labs(title = "count_N BA.2A and BA.2B")
```


Date -> numeric
```{r}
date <- as.numeric(covid_specific_1$date)
covid_specific_1$date <- date


r <- rep(NA,10000)

for (i in 1:length(r)) {
  j    <- sample(1:nrow(covid_specific_1), replace = TRUE)
  r[i] <- cor(covid_specific_1$date[j], covid_specific_1$count_N[j], method = "pearson")
  if(i%%500==0) {
    cat(i, "\n")
    flush.console()
  }
}

pd <- tibble(i=1:length(r), value=r)

ggplot(data = pd, mapping = aes(x=value)) + 
  geom_histogram(bins = 100) +
  labs(title = "Cor Coef Count_N XBB bs10000") +
  NULL
```
>Calculate the 95% bootstrapped confidence interval by getting the 2.5% and 97.5% percentile of the bootstrapped correlation coefficients

```{r}
quantile(x = r, probs = c(0.025, 0.975))
pd %>% 
  summarise(q_low  = quantile(value, probs = 0.025),
            q_high = quantile(value, probs = 0.975))
```

>Q: Compare it with the output of cor.test

```{r}
cor.test(covid_specific_1$date, covid_specific_1$count_N)
```

> By using QQ plot

```{r}
library(viridis)
# library(qqplotr)
ggplot(data = pd, aes(sample=value)) +
  # stat_qq_band(bandType = "boot", mapping = aes(fill = "Bootstrap"), alpha = 0.9) +
  #   stat_qq_line() +
  #   stat_qq_point() +
  # stat_qq_band(bandType = "ks", mapping = aes(fill = "KS"), alpha = 0.5) + ########### qqplotr has some bugs
  stat_qq(shape=21,fill="gray",size=4,colour="black",stroke=.5) +
  stat_qq_line(color="red") +
    labs(x = "Theoretical Quantiles", y = "Sample Quantiles", title = "QQ plot XBB") +
    scale_fill_viridis(discrete = T,direction = -1)
```

## hist of dnds
```{r}


tab <- c(1.184383, 0.4835135, 0.6762425, 0.3400288, 4.130618, 1.10106, 0.6724922, 1.566315, -13.42258 
) 
var <- c("BA.1",	"BA.2", "BA.2A", "BA.2B",	"BA.2.75" ,"BA.5"	,"BF.7","BQ.1.1", "XBB")
tb <- data.frame(var,tab)
tb$var <- factor(tb$var,levels = c("BA.1",	"BA.2", "BA.2A", "BA.2B",	"BA.2.75","BA.5"	,"BF.7","BQ.1.1", "XBB"))
# plot(tb, main = bquote('rdNdS'))

tb$col<- ifelse(tb$tab >= 1, 1, 0)
tb$col<-as.factor(tb$col)


ggplot(tb,mapping = aes(x = tb$var, y = tb$tab, col = tb$col, fill = tb$col)) +
  geom_bar(position="dodge",stat="identity") +
  labs(title = "dnds_new: kn/ks Bootstrapping") 
```

