---
title: "covid"
author: "zly"
date: '2022-07-25'
output: html_document
---

```{r}
library(rehh)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(data.table)
library(DBI)
library(RSQLite)

```

```{r}

filename <- file.choose()
covid <- readRDS(filename)

```

Modify on file
```{r}
# 
# # covid$country[covid$country == 'Taiwan'] <- 'Taiwan Province of China'
# 
# covid$VarShort[covid$Variant == 'VOC Omicron GRA (B.1.1.529+BA.*) first detected in Botswana/Hong Kong/South Africa']  <- 'VOC Omicron GRA'
# covid$VarShort[covid$Variant == 'VOC Alpha GRY (B.1.1.7+Q.*) first detected in the UK']  <- 'VOC Alpha GRY'
# covid$VarShort[covid$Variant == 'VOC Delta GK (B.1.617.2+AY.*) first detected in India']  <- 'VOC Delta GK'
# covid$VarShort[covid$Variant == 'VOC Beta GH/501Y.V2 (B.1.351+B.1.351.2+B.1.351.3) first detected in South Africa']  <- 'VOC Beta GH'
# covid$VarShort[covid$Variant == 'VOI Epsilon GH/452R.V1 (B.1.429+B.1.427) first detected in USA/California']  <- 'VOI Epsilon GH'
# covid$VarShort[covid$Variant == 'VOC Gamma GR/501Y.V3 (P.1+P.1.*) first detected in Brazil/Japan']  <- 'VOC Gamma GR'
# covid$VarShort[covid$Variant == 'VUM GH/490R (B.1.640+B.1.640.*) first detected in Congo/France']  <- 'VUM GH'
# covid$VarShort[covid$Variant == 'VOI Zeta GR/484K.V2 (P.2) first detected in Brazil']  <- 'VOI Zeta GR'
# covid$VarShort[covid$Variant == 'VOI Mu GH (B.1.621+B.1.621.1) first detected in Colombia']  <- 'VOI Mu GH'
# covid$VarShort[covid$Variant == 'VOI Iota GH/253G.V1 (B.1.526) first detected in USA/New York']  <- 'VOI Iota GH'
# covid$VarShort[covid$Variant == 'VOI Kappa G/452R.V3 (B.1.617.1) first detected in India']  <- 'VOI Kappa G'
# covid$VarShort[covid$Variant == 'VOI Lambda GR/452Q.V1 (C.37+C.37.1) first detected in Peru']  <- 'VOI Lambda GR'
# covid$VarShort[covid$Variant == 'VOI Eta G/484K.V3 (B.1.525) first detected in UK/Nigeria']  <- 'VOI Eta G'
# covid$VarShort[covid$Variant == 'VOI Theta GR/1092K.V1 (P.3) first detected in the Philippines']  <- 'VOI Theta GR'
# covid <- filter(covid,covid$n_total != 0)
# # colnames(covid)[names(covid) == 'AA Substitutions'] <- "AASubstitutions" 
# # saveRDS(covid, file = filename)



```

```{r}
# colnames(covid)[names(covid) == 'AA Substitutions'] <- "AASubstitutions" 
# colnames(covid)[names(covid) == 'GC-Content'] <- "GCContent"
# saveRDS(covid, file = filename)
```


1. Some plots based on !annotated
```{r}

set.seed(3373)
covid_small <- covid[sample(nrow(covid), 10000, replace = FALSE),]
ggplot(data = covid_small, mapping = aes(x = date, y = n_total/seq_length, col = region)) +
  geom_point() +
  xlab('date') +
  labs(title = "dnds, calculated and mean after Bootstrapping")

```

2. some plots based on annotated
```{r}
set.seed(3373)
covid_small <- covid[sample(nrow(covid), 20000, replace = FALSE),]
# Pr of 3 different mutations
dLength <- nrow(covid_small)
dNum <- 1

sub = 0
ins = 0
del = 0
while(dNum < dLength + 1){
  dV1 <- covid_small[dNum,3]
  dV2 <- covid_small[dNum,4]
  if(nchar(dV1) == nchar(dV2)){
    sub = sub + 1
  }
  else if(nchar(dV1) > nchar(dV2)){
    del = del + 1
  }
  else if(nchar(dV1) < nchar(dV2)){
    ins = ins + 1
  }
  dNum = dNum + 1
}
sub
del
ins

```

3. # of different types of variants based on !annotated
```{r}

barplot(table(covid_small$pangolin_lineage))

covid1 <- covid[covid$VarShort == c("VOC Delta GK"),] %>% filter(n_total > 50)

# ggplot(data = covid1, mapping = aes(x = date, y = n_total, col = covid1$country)) +
#   geom_point() +
#   xlab('date')

covid1 %>% count(country) %>% arrange(-n)

```


4. lineages
4.1 Print to file
```{r}
k <- unique(covid$pangolin_lineage)
write.table(k,file = "pangolin_lineage.txt", sep = " ", col.names = TRUE, row.names = FALSE)
```
4.2 By python using Trie

5. Variants of Concern
Omicron (B.1.1.529, BA.1, BA.1.1, BA.2, BA.3, BA.4 and BA.5 lineages)
```{r}

varofcon <- covid[covid$pangolin_lineage == c("B.1.1.529","BA.1","BA.2","BA.3","BA.4","BA.5"),]

pconcern <- ggplot(data = varofcon, mapping = aes(x = date, y = n_total/seq_length, col = pangolin_lineage)) +
  geom_point() +
  xlab('date')
pconcern <- pconcern + labs(title = "Variants of Concern")
pconcern

```
zly: intraspecific competition?


2022.9.18
6.synonmous based on !annotated
```{r}

# barplot(table(covid$pangolin_lineage))

ggplot(data = covid, mapping = aes(x = date, y = count_S, col = covid$VarShort)) +
  geom_point() +
  xlab('date') +
  ylim(0,50) +
  labs(title = "Synonmous accumulation in different variants by date")

```
Non-syn
```{r}
# barplot(table(covid$pangolin_lineage))

ggplot(data = covid, mapping = aes(x = date, y = count_N, col = covid$VarShort)) +
  geom_point() +
  xlab('date') +
  labs(title = "Non Synonmous accumulation in different variants by date") +
  ylim(0,100)
```

So most concerned should be: NA, VOC Alpha GRY, VOC Delta GK, VOC Omicron GRA

7. To get all the variants with AA Substitutions == "Spike_E484Q"
```{r}
# covid_small$id[which(grepl("Spike_E484Q",covid_small))]
covid_select_byE484Q <- covid %>%
  filter(grepl(pattern = 'Spike_E484Q', AASubstitutions))
```

```{r}
ggplot(data = covid_select_byE484Q, mapping = aes(x = date, y = count_S, col = covid_select_byE484Q$VarShort)) +
  geom_point() +
  xlab('date') +
  labs(title = "Spike_E484Q Synonmous")
```
Omicron spike mutations of interest: L452R, F486V, R493Q


8. computing the slopes
```{r}
d2 <- covid %>%
  filter(VarShort == 'VOC Alpha GRY')
m2 <- lm(count_S~date,d2)
summary(m2)

d3 <- covid %>%
  filter(VarShort == 'VOC Delta GK')
m3 <- lm(count_S~date,d3)
summary(m3)

d1 <- covid %>%
  filter(VarShort == 'VOC Omicron GRA')
m1 <- lm(count_S~date,d1)
summary(m1)
# plot(m1)


```
```{r}
covid <- covid[covid$VarShort == c("VOC Omicron GRA", "VOC Alpha GRY", "VOC Delta GK", "VOC Beta GH"),]
ggplot(data = covid, mapping = aes(x = date, y = count_S, col = covid$VarShort)) +
  # geom_point() +
  xlab('date') +
  ylim(0,20) +
  labs(title = "Mutation rate of different variants") +
  geom_smooth(method = lm)
```


9. GC content showing the stability
```{r}
covid$GCContent <- as.numeric(covid$GCContent)
ggplot(data = covid, mapping = aes(x = date, y = GCContent, col = covid$VarShort)) +
  geom_point() +
  xlab('date') +
  labs(title = "GC Content") +
  ylim(0.35,0.4)
```
The stability of RNA of different variants are not changed significantly, so the different mutation rate would come from other reasons.


10. Checking Num of guanine variaty
```{r}
library(gridExtra)
ggplot(data = covid, mapping = aes(x = date, y = G, col = covid$VarShort)) +
  geom_point() +
  labs(title = "# Guanine")
```

```{r}
covid %>% filter(gene == "S") %>% count(aa_change) %>% arrange(-n)
```
```{r}
covid %>% filter(gene == "S") %>% count(effect) %>% arrange(-n)
```

```{r}

ggplot(covid,mapping = aes(x = date, y = count_S)) +
  geom_point()
ggplot(covid,mapping = aes(x = date, y = count_N)) +
  geom_point()

```
```{r}
ggplot(covid1,mapping = aes(x = date, y = count_S)) +
  geom_point()
ggplot(covid1,mapping = aes(x = date, y = count_N)) +
  geom_point()

```



